
\begin{thebibliography}{1}

\bibitem{DNArepair} Alberts, B., Johnson, A., Lewis, J. et al. \textit{Molecular Biology of the Cell}. 4th edition. New York: Garland Science; 2002.

\bibitem{Alex1} Alexandrov, L.B. et al. 2013. Signatures of mutational processes in human cancer. \textit{Nature} \textbf{500}, 415-421.

\bibitem{Alex2} Alexandrov, L.B., Nik-Zainal, S., Wedge, D.C., Campbell, P.J. and Stratton, M.R. 2013. Deciphering signatures of mutational processes operative in human cancer. \textit{Cell Rep.} \textbf{3}, 246-259.

\bibitem{Alex3} Alexandrov, L.B., Jones, P.H., Wedge, D.C., Sale, J.E., Campbell, P.J., Nik-Zainal, S. and Stratton, M.R. 2015. Clock-like mutational processes in human somatic cells. \textit{Nature Genetics} \textbf{47}, 12, 1402-1407.

\bibitem{Alex4} Alexandrov, L.B. and Stratton, M.R. 2014. Mutational signatures: the patterns of somatic mutations hidden in cancer genomes. \textit{Curr. Opin. Genet. Dev.} \textbf{24}, 52-60.

\bibitem{Alex5} Alexandrov, L.B., Nik-Zainal, S., Siu, H.C., Suet, Y.L. and Stratton, M.R. 2015. A mutational signature in gastric cancer suggests therapeutic strategies. \textit{Nature Communications} \textbf{6}, 8683.

\bibitem{Cemgil} Cemgil, A.T. 2009. Bayesian Inference for Nonnegative Matrix Factorisation Models. \textit{Computational Intelligence and Neuroscience} 2009.

\bibitem{Denver} Denver, D.R., Feinberg, S., Estes, S., Thomas, W.K. and Lynch, M. 2005. Mutation Rates, Spectra and Hotspots in Mismatch Repair-Deficient Caenorhabditis Elegans. \textit{Genetics} \textbf{170}, 107-113.

\bibitem{Flibotte}  Flibotte, S. et al. 2010. Whole-genome profiling of mutagenesis in Caenorhabditis elegans. \textit{Genetics} \textbf{185}, 431-441.

\bibitem{COSMIC} Forbes, S.A. et al. 2014. COSMIC: exploring the world's knowledge of somatic mutations in human cancer. \textit{Nucl. Acids Res.} \textbf{43}, 805-811.

\bibitem{DNAdamagerepair} Friedberg, E.C., Walker, G.C. and Siede, W. \textit{DNA repair and mutagenesis}. Washington, D.C.: ASM Press; 1995.

\bibitem{Helleday} Helleday, T. et al. 2014. Mechanisms underlying mutational signatures in human cancers. \textit{Nat. Rev. Genet.} \textbf{15}, 585-598.

\bibitem{overdisp} Hinde, J. and Demetrio, C.G.B. 1998. Overdispersion: Models and estimation. \textit{Computational Statistics and Data Analysis} \textbf{27}, 151-170.

\bibitem{Hollstein} Hollstein, M., Alexandrov, L.B., Wild, C.P., Ardin, M. and Zavadil, J. 2016. Base changes in tumour DNA have the power to reveal the causes and evolution of cancer. \textit{Oncogene} (2016), 1-10.

\bibitem{Ivek} Ivek, I. 2014. Probabilistic Formulations of Nonnegative Matrix Factorization. \textit{Computer Science Seminars at Ruder Boskovic Institute}.

\bibitem{Lee} Lee, D.D., Seung, H.S. 2000 Algorithms for non-negative matrix factorization. In \textit{Advances in Neural Information Processing} \textbf{13} (Proc. NIPS*2000).

\bibitem{Meier1} Meier, B., Cooke, S.L., Weiss, J., Bailly, A.P., Alexandrov, L.B., Marshall, J., Raine, K., Maddison, M., Anderson, E., Stratton, M.R., Gartner, A., Campbell, P.J. 2014. C. elegans whole-genome sequencing reveals mutational signatures related to carcinogens and DNA repair deficiency. \textit{Genome Research} \textbf{24}, 1624-1636.

\bibitem{Meier2} Meier, B., Gartner, A. 2014. Having a direct look: Analysis of DNA damage and repair mechanisms by next generation sequencing. \textit{Experimental Cell Research} \textbf{329}, 35-41.

\bibitem{Wald} Molenberghs, G., and Verbeke, G. 2007. Likelihood Ratio, Score, and Wald Tests in a Constrained Parameter Space. \textit{The American Statistician} \textbf{61}, 22-27.

\bibitem{NZ} Nik-Zainal, S. et al. 2012. Mutational processes molding the genomes of 21 breast cancers. \textit{Cell} \textbf{149}, 979-993.

\bibitem{NZ2} Nik-Zainal, S. et al. 2016. Landscape of somatic mutations in 560 breast cancer whole-genome sequences. \textit{Nature} \textbf{534}, 47-54.

\bibitem{Poon1} Poon, S.L. et al. 2014. Mutation signatures of carcinogen exposure: genome-wide detection and new opportunities for cancer prevention. \textit{Genome Med.} \textbf{6}, 24.

\bibitem{Poon2} Poon, S.L., Huang, M.N. et al. 2015. Mutation signatures implicate aristolochic acid in bladder cancer development. \textit{Genome Medicine} \textbf{7(1)}, 38. 

\bibitem{p53} Rivlin, N., Brosh, R., Oren, M. and Rotter, V. 2011. Mutations in the p53 Tumor Suppressor Gene. \textit{Genes Cancer.} \textbf{2(4)}, 466-474.

\bibitem{Roberts} Roberts, S.A. and Gordenin, D.A. 2014. Hypoermutation in human cancer genomes: footprints and mechanisms. \textit{Nature Reviews} \textbf{14}, 786-800.

\bibitem{Segovia} Segovia, R., Tam, A.S., and Stirling, P.C. 2015. Dissecting genetic and environmental mutation signatures with model organisms. \textit{Trends Genet.} \textbf{31(8)}, 465-474.

\bibitem{Shlien} Shlien, A., Campbell, B.B., de Borja, R., Alexandrov, L.B., Merico, D., Wedge, D. et al. 2014. Combined hereditary and somatic mutations of replication error repair genes result in rapid onset of ultra-hypermutated cancers. \textit{Nature Genetics} \textbf{47}, 257-262.

\bibitem{caveman} Stephens, P.J., Tarpey, P.S. et al. 2012. The landscape of cancer genes and mutational processes in breast cancer. \textit{Nature} \textbf{486}, 400-404.

\bibitem{Stratton} Stratton, M.R., Campbell, P.J. and Futreal, P.A. 2009. The cancer genome. \textit{Nature} \textbf{458}, 719-724.

\bibitem{MMR} Supek, F. and Lehner, B. 2015. Differential DNA mismatch repair underlies mutation rate variation across the human genome. \textit{Nature} \textbf{521}, 81-84.

\bibitem{Teh} Teh, Y.W., Jordan, M.I., Beal, M.J. and Blei, D.M. 2006. Hierarchical Dirichlet Processes. \textit{Journal of the American Statistical Association} \textbf{101(476)}, 1566-1581

\bibitem{Tian} Tian et al. 2015. Computational methods and resources for the interpretation of genomic variants in cancer. \textit{BMC Genomics} \textbf{16(8)}.

\bibitem{KRAS} Wang, M.T. et al. 2015. K-Ras Promotes Tumorigenicity through Suppression of Non-canonical Wnt Signaling. \textit{Cell} \textbf{163(5)}, 1237-51.

\bibitem{Wu} Wu, S., Powers, S., Zhu, W. and Hannun, Y.A. 2016. Substantial contribution of extrinsic risk factors to cancer development. \textit{Nature} \textbf{569}, 43-47.

\bibitem{pindel} Ye K., Schulz M.H., Long Q., Apweiler R., Ning Z. 2009. Pindel: a pattern growth approach to detect break points of large deletions and medium sized insertions from paired-end short reads. \textit{Bioinformatics} \textbf{25(21)}, 2865-71..

\bibitem{BNB} Zhou, M. et al. 2012. Beta-Negative Binomial Process and Poisson Factor Analysis. \textit{International Conference on Artificial Intelligence and Statistics} (AISTATS2012), JMLR W\&CP, \textbf{22}, 1462-1471.

\end{thebibliography}